---
title: "Customer Personality Analysis"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction

We are a team focus on the detailed analysis of a company’s ideal customers. Our team strives to help company and business to modify products for customers with special needs. This dataset includes 27 features. 

## Variable Description

We can roughly categorized variable as 4 higher level group to describe the data: 

People(Age, marital status, kids number…)

Product(Amount of spend on each product category last 2 year…)

Promotion(number of deals using promotion)

Places(where did customer purchases, online or in store)

## Project Goal
Our goal is to perform clustering to summarize customer segments. We are achieving the goal through three steps: data cleaning, exploratory data analysis(EDA), and segmentation.

# Data Cleaning
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(naniar)
library(gridExtra)
library(lubridate)
```

```{r 1.1.1}
df <- read.delim("marketing_campaign.csv", stringsAsFactors = FALSE)
head(df)
```


```{r 1.1.2}
dim(df)
```

```{r 1.1.3}
summary(df)
```
## Missing Values
```{r 1.2.1}
# Counting the total number of missing values and each variables in the data
n_miss(df)
miss_var_summary(df)
```
Notice that income is the only variable with missing data problem, we will just drop the customer data point with missing data. 
```{r 1.2.2}
# Drop NA values
df_customers <- na.omit(df)
dim(df_customers)
```

## Outliers
The variable Year_Birth gives us the birth year of customers, which is not very intuitive. So we create a new variable Age. 
```{r 1.3.1}
# Creating a new variable Age from Year of Birth 
customers_unfilter <- df_customers %>% 
  mutate(Age = 2022 - Year_Birth)
customers_unfilter %>% 
  select(Age) %>% 
  arrange(desc(Age)) %>% 
  top_n(10)
# identify the outliers
g1=ggplot(customers_unfilter, aes(y = Age)) + geom_boxplot(outlier.colour = 'red') 
g2=ggplot(customers_unfilter, aes(y = Income)) + geom_boxplot(outlier.colour = 'red',fill='#FFDD00') 
g3=ggplot(customers_unfilter, aes(x = Age)) + geom_histogram(fill='black')
g4=ggplot(customers_unfilter, aes(x = Income)) + geom_histogram(fill='#FFDD00')
grid.arrange(g1, g2, g3, g4, ncol=2)
```

```{r 1.3.2}
# Max Age is > 100 
# Dropping outliers by setting a cap on Income and Age
customers <- customers_unfilter %>% 
  filter(Income < 600000 & Age < 90)
dim(customers)
```

```{r 1.3.3}
unique(customers$Marital_Status)
```

## Collapsing Features
Some features have too many unnecessary categories, we can convert to only two categories. 
```{r 1.4.1}
# Collapsing marital Status into two categories: Single & Taken

customers <- customers %>% 
  mutate(Marital_Status = 
           replace(Marital_Status,
                   Marital_Status == "Divorced" | Marital_Status == "Widow" | 
                     Marital_Status == "Alone" | Marital_Status == "Absurd" | 
                     Marital_Status == "YOLO", 
                   "Single"))
customers <- customers %>% 
  mutate(Marital_Status = 
           replace(Marital_Status,
                   Marital_Status == "Together" | Marital_Status == "Married", 
                   "Taken"))
```

```{r 1.4.2}
# Collapsing the Education into two Categories: graduate and non-graduate
unique(customers$Education)
customers <- customers %>% 
  mutate(Education = 
           replace(Education, Education == "Graduation"| Education == "PhD" | 
                     Education == "Master" | Education == "2n Cycle",
                   "graduate"))
customers <- customers %>% 
  mutate(Education = 
           replace(Education, Education == "Basic", "non-graduate"))
```

```{r 1.4.3}
# Converting them to factors
customers <- customers %>% 
  mutate(Marital_Status = as.factor(Marital_Status), 
         Education = as.factor(Education))
```

## Renaming Features
```{r 1.5.1}
# Renaming the Mnt_ features
customers <- customers %>% 
  rename(wines = MntWines, fruits = MntFruits, meat = MntMeatProducts, 
         fish = MntFishProducts, sweet = MntSweetProducts, gold = MntGoldProds, 
         webpurchase = NumWebPurchases, catalog = NumCatalogPurchases, 
         Store = NumStorePurchases, webvisit = NumWebVisitsMonth, 
         dealpurchase = NumDealsPurchases)
```

```{r 1.5.2}
# Creating a new variable:Total_spent
customers <- customers %>% 
  mutate(Total_spent = wines + fruits + meat + fish + sweet + gold) %>%
  mutate(Total_num = webpurchase + catalog + Store + dealpurchase)
```

```{r 1.5.3}
# Creating a new variable:kids
customers <- customers %>% 
  mutate(Kids = Kidhome + Teenhome)
```

```{r 1.5.4}
#Dropping some redundant features
customers <- customers %>% 
  select(- ID, - Year_Birth, - Dt_Customer, - Z_CostContact, 
         - Z_Revenue, - Kidhome, - Teenhome)
str(customers)
head(customers, n = 5)
```

# EDA
## 1. Marital Status
```{r 2.1.1}
ggplot(customers, aes(x=Marital_Status, fill=Marital_Status)) + 
  geom_bar() + ggtitle("Marital Status")
g1=ggplot(customers, mapping=aes(x=Education)) + 
  geom_bar(fill = "lightblue") + ggtitle("Education Level")
```

## 2. Education
```{r 2.1.2}
g2=ggplot(customers, aes(x=Education, y=Income)) + 
  geom_boxplot(fill = "pink") + ggtitle("Income vs. Education Level")
grid.arrange(g1, g2, ncol=2)
```

## 3. Income 
```{r 2.1.3}
g1=ggplot(customers, aes(x=Income, y=Total_spent)) + 
  geom_point(col='Sky Blue 3')+
  scale_x_continuous()+
  ggtitle("Total Amount Spent vs. Income")
g2=ggplot(customers, aes(x=Income, y=Total_num)) + 
  geom_point(col='Sky Blue 3')+
  scale_x_continuous()+
  ggtitle("Total Num vs. Income")
grid.arrange(g1, g2, ncol=2)
```

## 4. Products 
```{r 2.1.4}
g1 <- ggplot(customers, aes(x = wines)) +
  geom_histogram(fill = "Sky Blue 3", binwidth = 10)+
  scale_x_continuous(breaks = seq(0, 1600, 100)) +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))
g2 <- ggplot(customers, aes(x = fruits)) +
  geom_histogram(fill = "Sky Blue 3", binwidth = 10)+
  scale_x_continuous(breaks = seq(0, 250, 20))
g3 <- ggplot(customers, aes(x = meat)) +
  geom_histogram(fill = "Sky Blue 3", binwidth = 10)+
  scale_x_continuous(breaks = seq(0, 1800, 200)) +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))
g4 <- ggplot(customers, aes(x = fish)) +
  geom_histogram(fill = "Sky Blue 3", binwidth = 10)+
  scale_x_continuous(breaks = seq(0, 300, 50))
g5 <- ggplot(customers, aes(x = sweet)) +
  geom_histogram(fill = "Sky Blue 3", binwidth = 10)+
  scale_x_continuous(breaks = seq(0, 300, 20)) +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))
g6 <- ggplot(customers, aes(x = gold)) +
  geom_histogram(fill = "Sky Blue 3", binwidth = 10)+
  scale_x_continuous(breaks = seq(0, 400, 50)) +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))
g7 <- ggplot(customers, aes(x = Total_spent)) +
  geom_histogram(fill = "Sky Blue 3", binwidth = 10)+
  scale_x_continuous(breaks = seq(0, 3000, 100)) +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))
g8 <- ggplot(customers)+
  geom_density(aes(x = wines), color = "Red", fill = 0.7)+
  geom_density(aes(x = fruits), color = "Orange", fill = 0.7)+
  geom_density(aes(x = meat), color = "Blue", fill = 0.7)+
  geom_density(aes(x = fish), color = "Green", fill = 0.7)+
  geom_density(aes(x = sweet), color = "Black", fill = 0.7)+
  geom_density(aes(x = gold), color = "Sky Blue 3", fill = 0.7)+
  coord_cartesian(xlim = c(0, 500))+
  xlab("")

grid.arrange(g1, g2, g3, g4, g5, g6, g7, g8, ncol=4)
```

## 5. Correlation Plot
```{r 2.1.5}
library(corrplot)

#Getting correlation matrix 
cust_cor <- cor(customers[,3:26])
corrplot(cust_cor, 
         method = "color", 
         type = "upper")
         # order = "hclust", 
         # col = brewer.pal(n=10, name="RdYlBu"))
# png(file="corr.png", res=300, width=4500, height=4500)
# dev.off()
```
Here, based on the correlation plot, we could find that AcceptCmp*(1 if customer accepted the offer in the 1st campaign, 0 otherwise) are very much likely not to have any correlation with other variables, thus, we will remove these when we do the segementation. Secondly, there are some interesting findings. For example, the Income is highly negativly related to webvisit, while dealpurchase seems like the only variable to have the positive relationship with webvist. It could tell us people who like are not with high income usually will go to website to to some deal searching. And it is consistant result with our real world problem.

## Categorical Data
```{r 2.2}
# Encoding the categorical features to numeric
customers_copy <- customers

customers_copy <- customers_copy %>% 
  mutate(Education = case_when(
    Education == "graduate" ~ 1, Education == "non-graduate" ~ 0))
customers_copy <- customers_copy %>% 
  mutate(Marital_Status = case_when(
    Marital_Status == "Taken" ~ 1, Marital_Status == "Single" ~ 0))

str(customers_copy$Education)
str(customers_copy$Marital_Status)
dim(customers_copy)
```

```{r 2.3.1}
# glimpse(customers_copy)
```

```{r 2.3.2}
miss_var_summary(customers_copy)
```

# Segmentation
## Pre-processing Data
```{r 3.1}
library(caret)
# preprocessing the data, use numerical data to do dimension reduction
customers_copy_pre <- 
  preProcess(customers_copy[,c(3:15, 23:25)], method = c("center", "scale"))

# normalizing 
customers_copy <- predict(customers_copy_pre, customers_copy[,c(3:15, 23:25)])
summary(customers_copy)
```

## PCA: Dimensionality Reduction
We use PCA as the segmentation method in this part, it is an unsupervised learning method.

Firstly, we need to standardize some of the features for PCA: Centering and Scaling

We anticipate to have 2 or 3 specific customer groups for the results
```{r 3.2}
library(FactoMineR)
library(factoextra)

#Running a PCA.
customers_copy_pca <- PCA(customers_copy, graph = FALSE)

#Exploring PCA()

# Getting the summary of the pca
summary(customers_copy_pca)

#Getting the variance of the first 8 new dimensions
customers_copy_pca$eig[,2][1:8]

#Getting the cummulative variance
customers_copy_pca$eig[,3][1:8]

#Tracing variable contributions in customers_pca
customers_copy_pca$var$contrib
```


### Visualizing PCA
#### 1. Eigenvalues/variances vs. the Number of Dimensions


```{r 3.3.1}
# Plot the eigenvalues/variances against the number of dimensions
fviz_screeplot(customers_copy_pca, ncp=5)
```
In this plot, we could find that the first principal component will explain about 46% variance, second principal component is about 11% and the third principal component is about 6%. Thus, we will use first two principal component to do the following analysis.


#### 2. Variable Contributions


```{r 3.3.2}
#Creating a factor map for the variable contributions
fviz_pca_var(customers_copy_pca, 
             col.circle = 'black',
             col.var = "contrib", 
             gradient.cols = c("red", "green"), 
             repel = TRUE)
```

#### 3. Top 5 Variable Contributions

```{r 3.3.3}
#Creating a factor map for the top 5 variables with the highest contributions.
fviz_pca_var(customers_copy_pca, 
             select.var = list(contrib = 5), 
             col.var = "contrib",
             gradient.cols = c("green", "blue"), 
             repel = TRUE)
```


#### 4. Top 5 Variable Contributions in Barplot


```{r 3.3.4}
fviz_contrib(customers_copy_pca, choice = "var", axes = 1, top = 5)
```


#### 5. Graph of Individuals


```{r 3.3.5}
fviz_pca_ind(customers_copy_pca, col.ind="cos2") +
  scale_color_gradient2(low="blue", mid="white",high="red", midpoint=0.50)
```

#### 6. Biplot

```{r 3.3.6}
fviz_pca_biplot(customers_copy_pca, col.ind = 'dark grey')
```


#### 7. Hierarchical Cluster

```{r 3.3.7}
Distance <- dist(customers_copy, method = 'euclidean')
hplot<-hclust(Distance, method='complete')
plot(hplot, hang = -1, cex = 0.6)
```


## K-Means Clustering

```{r 3.4.1}
#The elbow method to find cluster number with kmean method
library(purrr)
tot_withinss <- map_dbl(1:10, function(k){
  model <- kmeans(x = customers_copy, centers = k)
  model$tot.withinss
})

elbow_df <- data.frame(
       k = 1:10,
       tot_withinss = tot_withinss)
head(elbow_df)
#plotting the elbow plot
ggplot(elbow_df, aes(k, tot_withinss)) + 
  geom_line(col='blue') + 
  scale_x_continuous(breaks = 1:10)+
  geom_vline(xintercept=3,linetype=2)
#plot elbow with package function
fviz_nbclust(customers_copy,kmeans,method="wss")+
  geom_vline(xintercept=3,linetype=2)
```
We can see that two way of the scree plot methods will plot the same result. And we will choose three as the cluster number.


```{r 3.4.2}
library(ggfortify); library(ggplot2)

set.seed(1313)
res_kmeans <- kmeans(customers_copy, centers = 3, nstart = 10)
# autoplot(res_kmeans, data=customers_copy)
fviz_cluster(res_kmeans, 
             customers_copy, 
             geom = "point",
             ellipse.type = "norm",
             repel = TRUE)
```


```{r 3.4.3}
set.seed(3)

#Building a k-means model with a k of 2
customers_md <- kmeans(customers_copy, center = 3)

#Extracting the vector of cluster assignment from the model
clust_customers <- as.factor(customers_md$cluster)

#Building the segment_customers dataframe
segment_customers <- mutate(customers_copy, cluster = clust_customers)

#Calculating the mean for each category
count(segment_customers, cluster)


#Adding the cluster variable to the original dataframe
customers <- customers %>% mutate(cluster = segment_customers$cluster)
head(customers, n = 3)
#confirming
count(customers, cluster)
```


### Segmenting Results

```{r 3.5}
g1=ggplot(data = customers, aes(x = cluster, y = Income, fill = cluster))+
  geom_boxplot() + 
  theme(legend.position = "none")

g2=ggplot(data = customers, aes(x = cluster, y = Recency, fill = cluster))+
  geom_boxplot() + 
  theme(legend.position = "none")

g111=ggplot(data = customers, aes(x = cluster, y = Age, fill = cluster))+
  geom_boxplot() + 
  theme(legend.position = "none")

g112=ggplot(data = customers, aes(x = cluster, y = Kids, fill = cluster))+
  geom_boxplot() + 
  theme(legend.position = "none")

g3=ggplot(data = customers, aes(x = cluster, y = wines, fill = cluster))+
  geom_boxplot() + 
  theme(legend.position = "none")

g4=ggplot(data = customers, aes(x = cluster, y = fruits, fill = cluster))+
  geom_boxplot() + 
  theme(legend.position = "none")

g5=ggplot(data = customers, aes(x = cluster, y = meat, fill = cluster))+
  geom_boxplot() + 
  theme(legend.position = "none")

g6=ggplot(data = customers, aes(x = cluster, y = fish, fill = cluster))+
  geom_boxplot() + 
  theme(legend.position = "none")

g7=ggplot(data = customers, aes(x = cluster, y = sweet, fill = cluster))+
  geom_boxplot() + 
  theme(legend.position = "none")

g8=ggplot(data = customers, aes(x = cluster, y = gold, fill = cluster))+
  geom_boxplot() + 
  theme(legend.position = "none")

g9=ggplot(data = customers, aes(x = cluster, y = dealpurchase, fill = cluster))+
  geom_boxplot() + 
  theme(legend.position = "none")

g10=ggplot(data = customers, aes(x = cluster, y = webpurchase, fill = cluster))+
  geom_boxplot() + 
  theme(legend.position = "none")

g11=ggplot(data = customers, aes(x = cluster, y = catalog, fill = cluster))+
  geom_boxplot() + 
  theme(legend.position = "none")

g12=ggplot(data = customers, aes(x = cluster, y = Store, fill = cluster))+
  geom_boxplot() + 
  theme(legend.position = "none")

g13=ggplot(data = customers, aes(x = cluster, y = Total_num, fill = cluster))+
  geom_boxplot() + 
  theme(legend.position = "none")

grid.arrange(g1, g2, g111, g112, ncol=2)
grid.arrange(g3, g4, g5, g6, g7, g8, ncol=3)
grid.arrange(g9, g10, g11, g12, g13, ncol=3)
```

# Conclusion

Basically, from the k-means clustering, we can separate customers into three groups with the following characteristics:

Cluster 1: 

Low purchasing power customer

No wine/meat

time gap between the going to market is large

have more kids

Cluster 2: 

Median purchasing power customer

prefer deal/web, but have some level of purchasing potential

love buying wines and gold

Cluster 3: 

High purchasing power customer

Prefer catalog/store

Purchasing more product

# Future steps

We will explore factor analysis which is one of the most populare analysis technique in market analysis